---
title: "Statistical Analysis of CO2 Emissions From Vehicles"
author: "Alex Zakrzeski, Aditya Nayak, Purvi Jain, and Tracie Robinson"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options: 
  markdown: 
    wrap: 72
---

```{=html}
<style type = "text/css">
body{ /* Normal  */
      font-size: 15px;
}
td {  /* Table  */
  font-size: 15px;
}
h1.title {
  font-size: 22px;
  color: Black;
}
h4.author {
font-size: 22px;
  color: Black;
}
h2 { /* Header 2 */
    font-size: 18px;
  color: DarkRed;
}
h3 { /* Header 3 */
  font-size: 18px;
  font-family: "Times New Roman", Times, serif;
  color: DarkBlue;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 12px;
}
</style>
```

## I. Introduction

### Background Information

Over the past three decades, climate change has been a subject of global
interest. Researchers have discovered numerous factors that contribute
to climate change, with CO2 emissions from vehicles being a noteworthy
instance. This year alone, experts forecast that CO2 emissions from
vehicles will exceed three billion metric tons. CO2 emissions from
vehicles are extremely harmful because they are a driving force for
increasing temperatures and rising sea levels. A solution to lessen the
adverse effects of CO2 emissions from vehicles is for manufacturers to
produce vehicles that emit relatively low levels of CO2 emissions. This
manufacturer initiative has been ongoing for years, with Toyota
releasing the world's first mass-produced hybrid vehicle in 1997. Still,
vehicle manufacturers need to implement more initiatives to decelerate
climate change. Finally, there is a significant urgency for new
strategies to be developed to further educate the public on how CO2
emissions from vehicles are accelerating climate change.

### Data Overview

For this analysis, a data set from the Canadian government's official
open data website was used. The data were collected over a seven-year
period and comprise 7,385 vehicles manufactured on three continents:
Asia, Europe, and North America. There are nine variables in the
dataset. Six of these variables are categorical; the other three are
numeric. Below is a summary of the variables used in this analysis.

-   Categorical Variables
    -   *Number of Cylinders:* The number of cylinders in the engine of
        a vehicle
    -   *Fuel Type:* The fuel type of the vehicle
    -   *Make:* The company that manufactured the vehicle
    -   *Model:* The model of the vehicle
    -   *Transmission:* The type of transmission for the vehicle
    -   *Vehicle Class:* The class of the vehicle by utility, capacity,
        and weight
-   Quantitative Variables
    -   *CO2 Emissions:* The tailpipe carbon dioxide emissions in grams
        per kilometer
    -   *Engine Size:* The engine's displacement in liters
    -   *Fuel Consumption:* The combined city/highway fuel consumption
        in liters per 100 km

### Research Question and Design

The research question for this analysis was: What is the best predictor
or combination of predictors for high CO2 emissions in vehicles? The
dependent variable in the research was "CO2 Emissions," and the
remaining eight variables were independent variables. The research
design encompassed numerous statistical techniques, including
descriptive statistics, correlation tests, and visualization. Overall,
the findings offer a clearer understanding of which characteristics of
vehicles are the most responsible for high CO2 emissions, therefore
making individuals more likely to choose to drive eco-friendly vehicles
to slow climate change.

```{r, include = FALSE}
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
```

```{r, include = FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(corrplot)
library(psych)
library(tibble)
library(gt)
```

## II. Data Preparation

### Data Cleaning

Prior to the exploratory data analysis, several data-cleaning steps were
taken, which included checking for missing values, manipulating strings,
renaming variables, and converting data types. The purpose of the data
cleaning was to enhance the quality and integrity of the data for the
next steps of the analysis.

```{r}
data = read_csv("co2_emissions_canada.csv")

glimpse(data)

nacount <- function(x) {
  data <- x %>%
    summarize(across(.fns = ~sum(is.na(.)))) %>%
    pivot_longer(cols = everything(), 
                 names_to = "variables", 
                 values_to = "na_count") %>%
    filter(na_count > 0)
}

miss_vals  <- nacount(data)

unique(data$Make)

data <- data %>%
  mutate(Make = str_to_title(data$Make),
         Make = str_replace(Make, "Bmw", "BMW"),
         Make = str_replace(Make, "Fiat", "FIAT"),
         Make = str_replace(Make, "Gmc", "GMC"),
         Make = str_replace(Make, "Mini", "MINI"), 
         Make = str_replace(Make, "Srt", "SRT")) 

unique(data$`Vehicle Class`)

data <- data %>%
  rename(Class = `Vehicle Class`) %>%
  mutate(Class = str_to_title(data$`Vehicle Class`),
         Class = str_replace(Class, "Van - Cargo", "Cargo Van"),
         Class = str_replace(Class, "Pickup Truck - Small", "Small Pickup Truck"),
         Class = str_replace(Class, "Suv - Small", "Small SUV"),
         Class = str_replace(Class, "Station Wagon - Small", "Small Station Wagon"),
         Class = str_replace(Class, "Van - Passenger", "Passenger Van"),
         Class = str_replace(Class, "Suv - Standard", "Standard SUV"),
         Class = str_replace(Class, "Pickup Truck - Standard", "Standard Pickup Truck"),
         Class = str_replace(Class, "Station Wagon - Mid-Size", "Mid-Size Station Wagon"))

data <- data %>%
  rename(Engine_Size = `Engine Size(L)`,
         Fuel_Type = `Fuel Type`,
         Fuel_Consumption = `Fuel Consumption Comb (L/100 km)`,
         Emissions = `CO2 Emissions(g/km)`) %>%
  mutate(Cylinders = factor(Cylinders))
```

### Data Wrangling

After the data was cleaned, several data-wrangling steps were taken,
which included creating new variables, grouping data, and aggregating
data. The purpose of data wrangling was to make the data more efficient
and insightful.

```{r}
unique(data$Make)

usa_make <- c("Buick", "Cadillac", "Chevrolet", "Chrysler", "Dodge", "Ford", 
              "GMC", "Jeep", "Lincoln", "Ram", "SRT")

eur_make <- c("Alfa Romeo", "Aston Martin", "Audi", "Bentley", "BMW","FIAT",
              "Jaguar", "Lamborghini", "Land Rover", "Maserati", "Mercedes-Benz",
              "MINI", "Porsche", "Rolls-Royce", "Smart", "Volkswagen", "Volvo")

asia_make <- c("Acura", "Honda", "Hyundai", "Infiniti", "Kia", "Lexus", "Mazda",
               "Mitsubishi", "Nissan", "Scion", "Subaru", "Toyota", "Genesis",
               "Bugatti")

data %>%
  select(Emissions) %>%
  filter(Emissions %in% c(250, 251))

data <- data %>%
  mutate(Location_Make = case_when(Make %in% usa_make ~ "USA",
                                   Make %in% eur_make ~ "EUR",
                                   Make %in% asia_make ~ "Asia"),
         Emissions_Lvl = if_else(Emissions >= mean(Emissions), 
                                 "Above Average",
                                 "Below Average"))

unique(data$Location_Make)

loc_lvl_count <- data %>%
  group_by(Location_Make, Emissions_Lvl) %>%
  summarize(Frequency = n())

loc_lvl_count <- loc_lvl_count %>%
  mutate(Location_Make = factor(Location_Make, levels = c("Asia", "USA", "EUR")))

unique(data$Class)

data %>%
 group_by(Class) %>%
 summarize(count = n())

class_avgs <- data %>%
  filter(Class == "Small Pickup Truck" | Class == "Standard Pickup Truck" |
         Class == "Small Station Wagon" | Class == "Mid-Size Station Wagon" |
         Class == "Small SUV" | Class == "Standard SUV") %>%
  mutate(Class1 = case_when(Class == "Small Pickup Truck" | 
                            Class == "Standard Pickup Truck" ~ "Pickup Truck",
                            Class == "Small Station Wagon" |
                            Class == "Mid-Size Station Wagon" ~ "Station Wagon",
                            Class == "Small SUV" |
                            Class == "Standard SUV" ~ "SUV"),
         Class2 = case_when(Class == "Small Pickup Truck" | 
                            Class == "Small Station Wagon" |
                            Class == "Small SUV" ~ "Small",
                            Class == "Standard Pickup Truck" |
                            Class == "Mid-Size Station Wagon" |
                            Class == "Standard SUV" ~ "Mid-Size"),
         Class2 = factor(Class2, levels = c("Small", "Mid-Size"))) %>%
  group_by(Class1, Class2) %>%
  summarize(avg_em = mean(Emissions))

unique(data$Fuel_Type)

ft_avgs <- data %>%
  filter(Fuel_Type != "N") %>%
  group_by(Fuel_Type) %>%
  summarize(avg_fc = round(mean(Fuel_Consumption), 2)) %>%
  arrange(desc(avg_fc)) %>%
  mutate(Fuel_Type = factor(Fuel_Type, levels = c("D", "X", "Z", "E"),
                            labels = c("Diesel Fuel", "Regular Gasoline",
                                       "Premium Gasoline", "Ethanol Fuel")))

mode1 <- data %>%
  group_by(Emissions) %>%
  count() %>%
  arrange(desc(n)) %>%
  head(1)

mode2 <- data %>%
  group_by(Engine_Size) %>%
  count() %>%
  arrange(desc(n)) %>%
  head(1)

corr_data1 <- data %>%
  select_if(is.numeric)

cm <- cor(corr_data1)

colnames(cm) <- c("Engine Size", "Fuel Consumption", "Emissions")

corr_data2 <- data %>%
  mutate(Cylinders = as.numeric(Cylinders),
         Emissions_Lvl = if_else(Emissions_Lvl == "Above Average", 1 , 0)) %>%
  select(Emissions, Cylinders, Emissions_Lvl)

emissions_outliers <- data %>%
  select(Emissions) %>%
  filter(Emissions > quantile(Emissions, 0.75) + 1.5 * IQR(Emissions) | 
         Emissions < quantile(Emissions, 0.25) - 1.5 * IQR(Emissions))

nooutliers_data <- data %>%
  mutate(Co2Em_Outliers = if_else(
    Emissions > quantile(Emissions, 0.75) + 1.5 * IQR(Emissions) | 
    Emissions < quantile(Emissions, 0.25) - 1.5 * IQR(Emissions), 1, 0)) %>%
  filter(Co2Em_Outliers == 0)
```

## III. Data Exploration

### Descriptive Statistics

The first step in this exploratory data analysis was to analyze the
descriptive statistics of the numeric variables that measure central
tendency and dispersion. For this analysis, the descriptive statistics
of the variables Engine Size and CO2 Emissions are especially noteworthy
and will be examined in further detail. The mean of CO2 emissions was
250.58, the median was 246, and the mode was 242. The mode of this
variable indicates that 242 is the most common value for a vehicle's CO2
emissions, with 85 vehicles in the data taking this value. The mean of
Engine Size was 3.16, the median was 3, and the mode was 2. The mean of
this variable indicates that the average engine size for the vehicles in
the data is 3.16, whereas the median displays that the middle point of
the distribution for engine size is 3. This middle point in the
distribution is also known as the second quartile, indicating that 50%
of the values for engine size are less than or equal to 3. A descriptive
statistics table for the numeric variables is presented below.

```{r, results = "asis", fig.align = "center"}
data %>%
  select_if(is.numeric) %>%
  rename(`Engine Size` = Engine_Size, `Fuel Consumption` = Fuel_Consumption,
         `CO2 Emissions` = Emissions) %>%
  describe() %>%
  select(n, mean, sd, median, min, max) %>%
  rename(N = n, Mean = mean, `St. Dev.` = sd, Median = median, Min = min, Max = max) %>%
  as_tibble(rownames = "Variable") %>%
gt() %>%
  cols_label(Variable = md("*Variable*"), N = md("*N*"), 
             Mean = md("*Mean*"),`St. Dev.` = md("*St. Dev.*"), 
             Median = md("*Median*"), Min = md("*Min*"), Max = md("*Max*")) %>%
  fmt_number(N, decimals = 0, sep_mark = ",") %>%
  fmt_number(Mean, decimals = 2) %>%
  fmt_number(`St. Dev.`, decimals = 2) %>%
  fmt_number(Median, drop_trailing_zeros = TRUE) %>%
  fmt_number(Max, drop_trailing_zeros = TRUE) %>%
  cols_align("center") 
```

### Stacked Bar Chart

All the vehicles in the data were produced in the United States, Europe,
or Asia. The stacked bar chart below displays the number of vehicles in
the data and the number of vehicles that have above- and below-average
CO2 emissions for each manufacturing location. The chart shows that
Europe manufactured the most vehicles, followed closely by the US, then
Asia. However, Asia is the manufacturing location with the largest
portion of vehicles that have below-average CO2 emissions levels.

```{r, fig.dim = c(5.75, 3.75), fig.align = "center"}
ggplot(data = loc_lvl_count, 
       mapping = aes(x = Location_Make, y = Frequency, fill = Emissions_Lvl,)) + 
  geom_bar(position = "stack", stat = "identity", color = "black", width = 0.5) +
  geom_text(aes(label = after_stat(y), group = Location_Make), 
            stat = 'summary', fun = sum, vjust = -1, size = 2.75) +
  scale_fill_manual(values = c("#c0c2c4", "#005288")) + 
  ylim(c(0,3000)) +
  labs(title = "Vehicles by Manufacturing Location and CO2 Emissions Level",
       x = "Manufacturing Location", y = "Vehicles") +
  guides(fill = guide_legend(title = "")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(size = 10),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9),
        legend.text = element_text(size = 8),
        legend.position = "bottom",
        legend.spacing.x = unit(0.25, "cm"))
```

### Grouped Bar Chart

The average CO2 emissions for certain vehicle classes (i.e., pickup
truck, station wagon, and SUV) and their respective sizes (mid-size and
small) was examined. A grouped bar chart displayed below shows the
average CO2 emissions for various combinations of vehicle class and
size. As anticipated, all the mid-size vehicles have larger CO2
emissions than their smaller counterparts. SUV was the vehicle class
with the largest difference in average CO2 emissions between its two
sizes.

```{r, fig.dim = c(5.75, 3.75), fig.align = "center"}
ggplot(data = class_avgs, mapping = aes(x = Class1, y = avg_em, fill = Class2)) +
  geom_col(position = "dodge", color = "black", width = 0.5) +
  ylim(0, 340) +
  geom_text(aes(label = round(avg_em, 2)), position = position_dodge(0.5),
            color = "black", vjust = -1, hjust = 0.5, size = 2.75) +
  scale_fill_manual(values = c("#c0c2c4", "#005288")) + 
  labs(title = "Average CO2 Emissions for Vehicles by Class and Size",
       x = "Class", y = "CO2 Emissions (g/km)") +
  guides(fill = guide_legend("")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(size = 10),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9),
        legend.text = element_text(size = 8),
        legend.position = "bottom",
        legend.spacing.x = unit(0.25, "cm"))
```

### Horizontal Bar Chart

In the data, a vehicle can have five different values for its fuel type.
Only one vehicle took natural gas; the most common fuel type was regular
gasoline. A horizontal bar chart to display the average fuel consumption
among the vehicles for the categories of fuel type except for natural
gas was created. Ethanol fuel had the highest average fuel consumption,
and diesel fuel had the lowest average. The horizontal bar chart supports
the preliminary research findings and is displayed below.

```{r, fig.dim = c(5.75, 3.75), fig.align = "center"}
ggplot(data = ft_avgs, mapping = aes(x = Fuel_Type, y = avg_fc)) +
  geom_col(color = "black", width = 0.5, fill = "#005288") +
  ylim(0, 18) +
  geom_text(aes(label = avg_fc), vjust = 0.5, hjust = -0.25, colour = "black", size = 2.75) +
  coord_flip() +
  labs(title ="Average Fuel Consumption by Fuel Type",
       x = "Fuel Type", y = "Fuel Consumption (L/100 km)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(size = 10),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9))
```

### Correlation Matrix

The Pearson correlation coefficients between the numeric variables in
the data were generated, which shows the linear strength of the
relationship between two continuous variables. A correlation matrix was
created to visually display this. All the numeric variables have a high 
positive correlation (\> 80%) with CO2 emissions. The analysis will further
investigate the relationship between a vehicle's engine size and its C02
emissions in this analysis. The correlation coefficient of these
variables indicates that as the engine size of a vehicle increases, its
level of CO2 emissions increases as well. This correlation matrix is
displayed below. Because Cylinders is an ordinal variable, the Spearman
rank correlation coefficient between Cylinders and CO2 Emissions was
generated. This correlation test measured the level of monotonicity
between the variables rather than the linearity. The correlation was
0.85, which indicates a strong nonlinear relationship.

```{r, fig.dim = c(8.5, 3.75), fig.align = "right"}
sp_cor <- cor(corr_data2$Cylinders, corr_data2$Emissions, method = "spearman")
corrplot.mixed(cm, lower.col = "black", number.cex = 0.6, tl.cex = 0.7, 
               mar = c(0, 0, 0.75, 3.5)) 
```

### Scatter Plot

Next, the analysis explored the relationship between a vehicle's engine
size and its C02 emissions in another data visualization, a scatter
plot. Engine Size is on the x-axis, and CO2 Emissions is on the y-axis.
The scatter plot displays a strong positive correlation. The linear
regression line displays the line of best fit for the data. Overall,
this linear regression line illustrates that the engine size of vehicles
is an excellent predictor of CO2 emissions. This scatter plot is
displayed below.

```{r, fig.dim = c(5.75, 3.75), fig.align = "center"}
ggplot(data = data, mapping = aes(x = Engine_Size, y = Emissions)) +
  geom_point(color = "#005288", size = 0.6) +
  geom_smooth(method = lm, se = FALSE, col = "black", size = 0.5) +
  labs(title = "CO2 Emissions vs. Engine Size",
       x = "Engine Size (L)", y = "CO2 Emissions (g/km)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(size = 10),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9))
```

### Density Plots

The distribution of CO2 emissions is examined by creating a kernel
density plot. The vertical line displays the median value for CO2
Emissions. As seen in the plot, the CO2 Emissions variable is normally
distributed. This was anticipated because the variable's mean, median,
and mode are all extremely close. This density plot is displayed below.

```{r, fig.dim = c(5.75, 3.75), fig.align = "center"}
ggplot(data = data, mapping = aes(x = Emissions)) +
  geom_density(fill = "#c0c2c4", alpha = 0.7) +
  geom_vline(xintercept = median(data$Emissions), linetype = "dotted") +
  scale_y_continuous(labels = function(x) paste0(x * 100, "%")) +
  labs(title = "Distribution of CO2 Emissions", x = "CO2 Emissions (g/km)", y = "Density") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(size = 10),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9))
```

The distribution of Engine Size is examined by creating a kernel density
plot. The vertical line displays the median value for Engine Size. The
distribution displays a right skew. This was anticipated this because
the mean of the Engine Size is larger than its median. Finally, the
distribution appears to be multimodal. This density plot is displayed
below.

```{r, fig.dim = c(5.75, 3.75), fig.align = "center"}
ggplot(data = data, mapping = aes(x = Engine_Size)) +
  geom_density(fill = "#005288", alpha = 0.7) +
  geom_vline(xintercept = median(data$Engine_Size), linetype = "dotted") +
  scale_y_continuous(labels = function(x) paste0(x * 100, "%")) +
  labs(title = "Distribution of Engine Size", x = "Engine Size (L)", y = "Density") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(size = 10),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9))
```

### Box Plot

A box plot was created to display the distribution of CO2 Emissions. It
is apparent in the visualization that there are numerous outliers when
looking at the distribution of the variable. These outliers can be seen
outside the upper whisker of the box plot. The interquartile range
method was used to remove these outliers from the distribution of the
variable. Ultimately, eighty outliers were removed from the distribution
of CO2 Emissions variable, and all the rows that contained outliers for
the variable were dropped.

```{r, fig.dim = c(5.75, 3.75), fig.align = "center"}
ggplot(data = data, mapping = aes(x = "", y = Emissions)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(data = emissions_outliers, color = "#005288", height = 0, width = 0.1) +
   labs(title = "Distribution of CO2 Emissions",
        x = "Vehicles", y = "CO2 Emissions (g/km)") +
  theme_bw() +
   theme(plot.title = element_text(hjust = 0.5, size = 11),
        axis.title = element_text(size = 10),
        axis.text.x = element_text(size = 9),
        axis.text.y = element_text(size = 9))
```

## IV. Conclusion

### Overview of Findings

Many noteworthy findings emerged from this statistical analysis of the
relationship between vehicles' CO2 emission and their characteristics.
The exploratory data analysis employed descriptive statistics,
visualization, and correlation tests. A significant finding was the
strong linear relationship between vehicles' CO2 emissions and engine
size. This was shown through a high Pearson correlation coefficient of
0.85. This relationship was also displayed on a scatter plot with a
linear regression line.

Another insightful finding was just as strong of a correlation existed
between CO2 emissions and engine cylinders. Because Cylinders was an ordinal
variable, the Spearman correlation coefficient between Cylinders and CO2
Emissions was generated, which was 0.85. This relationship is nonlinear,
unlike that of CO2 emissions and engine size.
 
Additional analyses can come from numerous statistical models. Linear
regression and k-nearest neighbors are two algorithms that can be
further explored in future work that encompass predictive models.
Overall, the results back the majority of the initial research findings,
and individuals can better understand what characteristics of vehicles
contribute to the release of excessive CO2 emissions into the earth's
atmosphere.

::: {.tocify-extend-page data-unique="tocify-extend-page" style="height: 5px;"}
:::
