---
title: "CO2 Emissions Data Analysis"
author: "Alexander Zakrzeski"
date: "October 8th, 2022"
output: word_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

These are the packages used in this data analysis.

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(corrplot)
library(psych)
library(AICcmodavg)
library(fastDummies)
library(caret)
```

I. Data Overview and Cleaning

The data set is from the official open data web site of the Canadian government.
It depicts numerous details of how the CO2 emissions for vehicles can differ based 
on certain characteristics. The data was collected over a seven year period, and
there are 7,385 vehicles in this data set. Different vehicles with all of the same 
characteristics can appear more than once in the data set.

First, import the data. Then, check for missing values in the data. No missing 
values were found.

```{r}
data = read_csv("co2_emissions_canada.csv")

glimpse(data)

nacount <- function(x) {
  data <- x %>%
    summarize(across(.fns = ~sum(is.na(.)))) %>%
    pivot_longer(cols = everything(), 
                 names_to = "variables", 
                 values_to = "na_count") %>%
    filter(na_count > 0)
}

miss_vals  <- nacount(data)
```

After checking for missing values, more data cleaning steps are taken. These steps
include the following: string manipulation and renaming variables.

```{r}
unique(data$Make)

data <- data %>%
  mutate(Make = str_to_title(data$Make),
         Make = str_replace(Make, "Bmw", "BMW"),
         Make = str_replace(Make, "Fiat", "FIAT"),
         Make = str_replace(Make, "Gmc", "GMC"),
         Make = str_replace(Make, "Mini", "MINI"), 
         Make = str_replace(Make, "Srt", "SRT")) 

unique(data$`Vehicle Class`)

data <- data %>%
  rename(Class = `Vehicle Class`) %>%
  mutate(Class = str_to_title(data$`Vehicle Class`),
         Class = str_replace(Class, "Van - Cargo", "Cargo Van"),
         Class = str_replace(Class, "Pickup Truck - Small", "Small Pickup Truck"),
         Class = str_replace(Class, "Suv - Small", "Small SUV"),
         Class = str_replace(Class, "Station Wagon - Small", "Small Station Wagon"),
         Class = str_replace(Class, "Van - Passenger", "Passenger Van"),
         Class = str_replace(Class, "Suv - Standard", "Standard SUV"),
         Class = str_replace(Class, "Pickup Truck - Standard", "Standard Pickup Truck"),
         Class = str_replace(Class, "Station Wagon - Mid-Size", "Mid-Size Station Wagon"))

data <- data %>%
  rename(Engine_Size = `Engine Size(L)`,
         Fuel_Type = `Fuel Type`,
         Fuel_Consumption = `Fuel Consumption Comb (L/100 km)`,
         Emissions = `CO2 Emissions(g/km)`)
```

II. Data Exploration

The next step in this analysis is exploring the descriptive statistics of all the 
numeric variables in the data set. A descriptive statistics table is created that 
looks at both central tendency and dispersion. There are many noteworthy findings
when looking at the descriptive statistics. For example, the average CO2 emissions 
for all vehicles is 250.58 g/km. Finally, a frequency distribution table is created
for the variable that represents the number cylinders in a vehicle's engine because
it will be treated as a categorical variable for certain parts of this analysis.
In the data, cars with four-cylinder engines are the most common.

```{r}
ds <- data %>%
  select_if(is.numeric) %>%
  select(-Cylinders) %>%
  rename(`Engine Size` = Engine_Size, `Fuel Consumption` = Fuel_Consumption) %>%
  describe() %>%
  select(n, mean, sd, median, min, max) %>%
  rename(N = n, Mean = mean, `St. Dev.` = sd, Median = median, Min = min, Max = max)

print(ds)

fd <- data %>%
  group_by(Cylinders) %>%
  summarize(Frequency = n())

print(fd)
```

All of the vehicles in the data are produced either in the United States, Europe
or Asia. As stated before, the mean of CO2 emissions for all the vehicles is 
250.8 g/km. Looking at the values of CO2 emissions in the data, if the variable
has a value above 250.8 it is above average. Thus, when the value is below 250.8
it is below average. Also, it is important to note that the variable for C02
emissions has no values equal to the mean. A stacked bar chart is created that
displays the number of vehicles in the data that are manufactured in the United States,
Europe and Asia as well as the number of vehicles that have above and below average
CO2 emissions for each manufacturing location.

```{r}
unique(data$Make)

usa_make <- c("Buick", "Cadillac", "Chevrolet", "Chrysler", "Dodge", "Ford", 
              "GMC", "Jeep", "Lincoln", "Ram", "SRT")

eur_make <- c("Alfa Romeo", "Aston Martin", "Audi", "Bentley", "BMW","FIAT",
              "Jaguar", "Lamborghini", "Land Rover", "Maserati", "Mercedes-Benz",
              "MINI", "Porsche", "Rolls-Royce", "Smart", "Volkswagen", "Volvo")

asia_make <- c("Acura", "Honda", "Hyundai", "Infiniti", "Kia", "Lexus", "Mazda",
               "Mitsubishi", "Nissan", "Scion", "Subaru", "Toyota", "Genesis",
               "Bugatti")

data %>%
  select(Emissions) %>%
  filter(Emissions %in% c(250, 251))

data <- data %>%
  mutate(Location_Make = case_when(Make %in% usa_make ~ "USA",
                                   Make %in% eur_make ~ "EUR",
                                   Make %in% asia_make ~ "Asia"),
         Emissions_Lvl = if_else(Emissions >= mean(Emissions), 
                                 "Above Average",
                                 "Below Average"))

unique(data$Location_Make)

loc_lvl_count <- data %>%
  group_by(Location_Make, Emissions_Lvl) %>%
  summarize(Frequency = n())

loc_lvl_count <- loc_lvl_count %>%
  mutate(Location_Make = factor(Location_Make, levels = c("Asia", "USA", "EUR")))

ggplot(data = loc_lvl_count, 
       mapping = aes(x = Location_Make, y = Frequency, fill = Emissions_Lvl,)) + 
  geom_bar(position = "stack", stat = "identity", color = "black", width = 0.5) +
  geom_text(aes(label = after_stat(y), group = Location_Make), 
            stat = 'summary', fun = sum, vjust = -1, size = 3) +
  scale_fill_manual(values = c("#1170aa", "#e03531")) + 
  ylim(c(0,3000)) +
  labs(title = "Number of Vehicles by Manufacturing Location and CO2 Emissions Level",
       x = "Manufacturing Location", y = "Vehicles") +
  guides(fill = guide_legend(title = "CO2 Emissions Level")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8))
```

Now the average CO2 emissions for certain vehicle classes and their respective
sizes will be examined. The three vehicle classes included are the following:
pickup truck, station wagon and SUV. Then, the two vehicle sizes included are the
following: mid-size and small. A grouped bar chart displays the average CO2 
emissions for the various combinations of vehicle class and size. As anticipated,
all the mid-size vehicles have larger CO2 emissions than their smaller counterparts.
SUV was the vehicle class that has largest difference in average CO2 emissions
between its two sizes.

```{r}
unique(data$Class)

data %>%
 group_by(Class) %>%
 summarize(count = n())

class_avgs <- data %>%
  filter(Class == "Small Pickup Truck" | Class == "Standard Pickup Truck" |
         Class == "Small Station Wagon" | Class == "Mid-Size Station Wagon" |
         Class == "Small SUV" | Class == "Standard SUV") %>%
  mutate(Class1 = case_when(Class == "Small Pickup Truck" | 
                            Class == "Standard Pickup Truck" ~ "Pickup Truck",
                            Class == "Small Station Wagon" |
                            Class == "Mid-Size Station Wagon" ~ "Station Wagon",
                            Class == "Small SUV" |
                            Class == "Standard SUV" ~ "SUV"),
         Class2 = case_when(Class == "Small Pickup Truck" | 
                            Class == "Small Station Wagon" |
                            Class == "Small SUV" ~ "Small",
                            Class == "Standard Pickup Truck" |
                            Class == "Mid-Size Station Wagon" |
                            Class == "Standard SUV" ~ "Mid-Size")) %>%
  group_by(Class1, Class2) %>%
  summarize(avg_em = mean(Emissions))

ggplot(data = class_avgs, mapping = aes(x = Class1, y = avg_em, fill = Class2)) +
  geom_col(position = "dodge", color = "black", width = 0.5) +
  ylim(0, 340) +
  geom_text(aes(label = round(avg_em, 2)), position = position_dodge(0.5),
            color = "black", vjust = -1, hjust = 0.5, size = 2.5) +
  scale_fill_manual(values = c("#1170aa", "#e03531")) + 
  labs(title = "Average CO2 Emissions of Vehicles by Class and Size",
       x = "Class", y = "CO2 Emissions (g/km)") +
  guides(fill = guide_legend("Size")) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
        axis.title = element_text(size = 8),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 8))
```

Next, the average fuel consumption for the different fuel types of the vehicles 
will be explored. In the data, a vehicle can have five different values for its
fuel type. Only one vehicle takes natural gas as its fuel type, while the most
common fuel type for the vehicles is regular gasoline. A horizontal bar chart
is created displaying the the average fuel consumption among the vehicles for all
the categories of fuel type except for natural gas. The preliminary research
findings back what the horizontal bar chart displays.

```{r}
unique(data$Fuel_Type)

ft_avgs <- data %>%
  filter(Fuel_Type != "N") %>%
  group_by(Fuel_Type) %>%
  summarize(avg_fc = round(mean(Fuel_Consumption), 2)) %>%
  arrange(desc(avg_fc)) %>%
  mutate(Fuel_Type = factor(Fuel_Type, levels = c("D", "X", "Z", "E"),
                            labels = c("Diesel Fuel", "Regular Gasoline",
                                       "Premium Gasoline", "Ethanol Fuel")))

ggplot(data = ft_avgs, mapping = aes(x = Fuel_Type, y = avg_fc)) +
  geom_col(color = "black", width = 0.5, fill = "#1170aa") +
  ylim(0, 18) +
  geom_text(aes(label = avg_fc), vjust = 0.5, hjust = -0.25, colour = "black", size = 3) +
  coord_flip() +
  labs(title ="Average Fuel Consumption by Fuel Type",
       x = "Fuel Type", y = "Fuel Consumption (L/100 km)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
        axis.title = element_text(size = 8))
```

Now the distribution for CO2 emissions will be visualized and examined through
creating a kernel density plot. The mean of CO2 emissions is 250.58, then the median
is 246, and finally the mode is 242. The mode of this variable indicates that 242
is the most common value for a vehicle's CO2 emissions with 85 vehicles in the 
data taking this particular value. As seen in the visualization, the distribution
of CO2 emissions is quite normally distributed. This was anticipated when looking
at the descriptive statistics because the mean, median and mode of the variable
are all extremely close to one another.

```{r}
mode1 <- data %>%
  group_by(Emissions) %>%
  count() %>%
  arrange(desc(n)) %>%
  head(1)

ggplot(data = data, mapping = aes(x = Emissions)) +
  geom_density(fill = "#1170aa", alpha = 0.7) +
  scale_y_continuous(labels = function(x) paste0(x * 100, "%")) +
  labs(title = "Distribution of CO2 Emissions", x = "CO2 Emissions (g/km)", y = "Density") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
        axis.title = element_text(size = 8))
```

Then, the distribution for engine size will each be visualized and examined through
creating a kernel density plot. The mean of engine size is 3.16, then the median
is 3, and finally the mode is 2. The mean of this variable indicates that the 
average engine size for the vehicles in the data is 3.16, while the median displays
that the middle point of the distribution for engine size is 3.This middle point
in the distribution is also known as the second quartile, and it indicates that
fifty percent of the values for engine size are less than or equal to 3. The 
distribution of the variable engine displays a right skew. This was anticipated
when looking at the descriptive statistics because the mean for engine size is
larger than its median. Finally, the distribution of this variable appears to be
multimodal.

```{r}
mode2 <- data %>%
  group_by(Engine_Size) %>%
  count() %>%
  arrange(desc(n)) %>%
  head(1)

ggplot(data = data, mapping = aes(x = Engine_Size)) +
  geom_density(fill = "#e03531", alpha = 0.7) +
  scale_y_continuous(labels = function(x) paste0(x * 100, "%")) +
  labs(title = "Distribution of Engine Size", x = "Engine Size (L)", y = "Density") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
        axis.title = element_text(size = 8))
```

The correlation between the numeric variables in the data are now examined. 
A correlation matrix is created to visually display this. The variable 
representing the number of cylinders in a vehicle's engine will be included in this 
visualization. All of the numeric variables have a high positive correlation
(>80%) with CO2 emissions. The relationship between the engine size of a vehicle
and its C02 emissions will be further investigated in this analysis. The correlation
coefficient of these variables indicates that as the engine size of a vehicle
increases there is a high likelihood that its level of CO2 emissions will increase
as well.

```{r}
corr_data <- data %>%
  select_if(is.numeric)

cm <- cor(corr_data)

colnames(cm) <- c("Engine Size", "Cylinders", "Fuel Consumption", "Emissions")

corrplot.mixed(cm, lower.col = "black", number.cex = 0.7, tl.cex = 0.5, 
               mar = c(0, 0, 0.75, 3.5)) 
mtext("Correlation Matrix of Numeric Variables", at = 2.5, line = 1.2, 
       cex = 0.8, font = 2)
```

Next, the relationship between the engine size of a vehicle and its C02 emissions is
explored further in another data visualization. A scatter plot is created to display the 
relationship between these two variables. Engine size is on the x-axis, while CO2
emissions is on the y-axis. The scatter plot displays the visual representation of
the strong positive correlation between the two variables. There is a linear 
regression line in the scatter plot that displays the line of best fit for the data.
Overall, this linear regression line in the scatter plot illustrates that the 
engine size of vehicles is an excellent predictor for their the level of CO2 
emissions.

```{r warning = FALSE, message = FALSE}
ggplot(data = data, mapping = aes(x = Engine_Size, y = Emissions)) +
  geom_point(color = "#1170aa", size = 0.6) +
  geom_smooth(method = lm, se = FALSE, col = "black", size = 0.5) +
  labs(title = "CO2 Emissions vs. Engine Size",
       x = "Engine Size (L)", y = "CO2 Emissions (g/km)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
        axis.title = element_text(size = 8))
```

A box plot is created to display the distribution of CO2 emissions for the vehicles.
It is apparent in the visualization that there are numerous outliers when looking
at the distribution of the variable. These outliers can be seen outside the upper
whisker of the box plot. The interquartile range method is used to remove these
outliers from the distribution of the variable. Ultimately, eighty outliers are
removed from distribution of CO2 emissions, and all the rows that contain outliers
for the variable are dropped.

```{r}
emissions_outliers <- data %>%
  select(Emissions) %>%
  filter(Emissions > quantile(Emissions, 0.75) + 1.5 * IQR(Emissions) | 
         Emissions < quantile(Emissions, 0.25) - 1.5 * IQR(Emissions))

ggplot(data = data, mapping = aes(x = "", y = Emissions)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(data = emissions_outliers, color = "#1170aa", height = 0, width = 0.1) +
   labs(title = "Distribution of CO2 Emissions",
        x = "Vehicles", y = "CO2 Emissions (g/km)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 10, face = "bold"),
        axis.title = element_text(size = 8))

nooutliers_data <- data %>%
  mutate(Co2Em_Outliers = if_else(
    Emissions > quantile(Emissions, 0.75) + 1.5 * IQR(Emissions) | 
    Emissions < quantile(Emissions, 0.25) - 1.5 * IQR(Emissions), 1, 0)) %>%
  filter(Co2Em_Outliers == 0)
```
